The importance of reproducibility in COVID-19 research: the case of population density and the spread of the pandemic

### 2021-05-19 R0-Density-Reanalysis/R0-Density-Reanalysis.tex
Added reference to butress the point of risk perception

A plausible behavioral adaptiation during a pandemic, especially one
broadcast as widely and intensely as COVID-19, is risk compensation.
Risk compensation is a process whereby people adjust their behavior in
response to their \emph{perception} of risk {[}28--30{]}. For example,
it is possible that people who listened to the message of leaders saying
that they were safe because of low density may not have taken adequate
precautions. People in dense places who could more directly observe the
impact of the pandemic may have become overly cautious. Both Paez et al.
{[}31{]} and Hamidi et al. {[}32{]} posit this mechanism (i.e., greater
compliance with social distancing in denser regions) to explain the
results of their analyses. The evidence available does indeed show that
there were important changes in behavior during the pandemic, at least
with respect to mobility {[}33--35{]}; furthermore, shelter in place
orders may have had greater buy-in from the public in higher density
regions {[}36{]}, and the behavior may have persisted beyond the
duration of official social-distancing policies {[}37{]}. In addition,
there is evidence that changes in mobility correlated with the
trajectory of the pandemic {[}38,39{]}. Given the potential for
behavioral adaptation, the question of density becomes more nuanced: it
is not just a matter of proximity, but also of human behavior, which is
better studied using population-level data and models.

@@

On the one hand, there are studies that report positive associations
between population density and various COVID-19-related outcomes. Bhadra
{[}40{]}, for example, reported a moderate positive correlation between
the spread of COVID-19 and population density at the district level in
India, however their analysis was bivariate and did not control for
other variables, such as income. Similarly, Kadi and Khelfaoui {[}41{]}
found a positive and significant correlation between number of cases and
population density in cities in Algeria in a series of simple regression
models (i.e., without other controls). A question in these relatively
simple analyses is whether density is not a proxy for other factors.
Other studies have included controls, such as Pequeno et al. {[}42{]}, a
team that reported a positive association between density and cumulative
counts of confirmed COVID-19 cases in state capitals in Brazil after
controlling for covariates, including income, transport connectivity,
and economic status. In a similar vein, Fielding-Miller et al. {[}43{]}
reported a positive relationship between the absolute number of COVID-19
deaths and population density (rate) in rural counties in the US. Roy
and Ghosh {[}44{]} used a battery of machine learning techniques to find
discriminatory factors, and a positive and significant association
between COVID-19 infection and death rates in US states. Wong and Li
{[}45{]} also found a positive and significant association between
population density and number of confirmed COVID-19 cases in US
counties, using both univariate and multivariate regressions with
spatial effects. Most recently, Sy et al. {[}19{]} reported that the
basic reproductive number of COVID-19 in US counties tends to increase
with population density, but at a decreasing rate at higher densities.

On the gripping hand, a number of studies report non-significant or
negative associations between population density and COVID-19 outcomes.
This includes the research of Sun et al. {[}46{]} who did not find
evidence of significant correlation between population density and
confirmed number of cases per day \emph{in conditions of lockdown} in
China. This finding echoes the results of Paez et al. {[}31{]}, who in
their study of provinces in Spain reported non-significant associations
between population density and infection rates in the early days of the
first wave of COVID-19, and negative significant associations in the
later part of the first lockdown. Similarly, {[}47{]} found zero or
negative associations between population density and infection
numbers/deaths by country. Fielding-Miller et al. {[}43{]} found a
negative relationship between COVID-19 deaths and population density
urban counties in the US. In their investigation of doubling time, White
and Hébert-Dufresne {[}48{]} identified a negative and significant
correlation between population density and doubling time in US states.
Likewise, {[}49{]} fond a small negative (and significant) association
between population density and COVID-19 morbidity in districts in
Tehran. And two of the most complete studies in the US by Hamidi et al.
{[}32,50{]} used an extensive set of controls to find negative and
significant correlations between density at the level of counties in the
US and COVID-19 cases and fatalities.

@@

There are various reasons why a project can fail to be reproducible. In
some cases, there might be legitimate reasons to withhold the data,
perhaps due to confidentiality and privacy reasons {[}e.g., 51{]}. But
in many other cases the data are publicly available, as has been
commonly the case with population-level COVID-19 information. Often the
provenance of the data is documented, but the data themselves are not
shared {[}25,32,40,43,50,e.g., 52,53--55{]}. As any researcher can
attest, whether a graduate student or a seasoned scientist, collecting,
organizing, and preparing data for a project can take a substantial
amount of time. Pointing to the sources of data, even when these sources
are public, is a small step towards reproducibility-but only a very
small one. Faced with the prospect of having to recreate a data set from
raw sources is probably sufficient to dissuade all but the most
dedicated (or stubborn) researcher. This is true even if part of the
data are shared {[}e.g., 45{]}. In other cases, data are shared, but the
processes to document the preparation of the data are not fully
documented {[}47,e.g., 56{]}. These processes matter, as shown by the
errors in the spreadsheets of Reinhart and Rogoff {[}16{]}, and the data
of biologist Jonathan Pruitt that led to an ``avalanche'' of paper
retractions\footnote{\url{https://doi.org/10.1038/d41586-020-00287-y}}.
Another situation is when papers share well-documented data, but fail to
provide the code used in the analysis {[}42,e.g., 57,58{]}. Making code
available only ``on demand'' {[}e.g., 59{]} is an unnecessary barrier
when most journals offer the facility to share supplemental materials
online. Then there are those papers that strive towards reproducibility,
sharing well-documented processes and data, as well as the code used in
any analyses reported {[}e.g., 31,36,48,60{]}.

@@

\leavevmode\hypertarget{ref-Paez2020spatio}{}%
\CSLLeftMargin{31. }
\CSLRightInline{Paez A, Lopez FA, Menezes T, Cavalcanti R, Pitta MG da
R. A spatio-temporal analysis of the environmental correlates of
COVID-19 incidence in spain. Geographical Analysis. 2020;n/a.
doi:\href{https://doi.org/10.1111/gean.12241}{10.1111/gean.12241}}

\leavevmode\hypertarget{ref-Hamidi2020density}{}%
\CSLLeftMargin{32. }
\CSLRightInline{Hamidi S, Sabouri S, Ewing R. Does density aggravate the
COVID-19 pandemic? Journal of the American Planning Association.
2020;86: 495--509.
doi:\href{https://doi.org/10.1080/01944363.2020.1777891}{10.1080/01944363.2020.1777891}}

\leavevmode\hypertarget{ref-Jamal2020Changes}{}%
\CSLLeftMargin{33. }
\CSLRightInline{Jamal S, Paez A. Changes in trip-making frequency by
mode during COVID-19. Findings. Network Design Lab; 2020;
doi:\href{https://doi.org/10.32866/001c.17977}{10.32866/001c.17977}}

\leavevmode\hypertarget{ref-Harris2021Changes}{}%
\CSLLeftMargin{34. }
\CSLRightInline{Harris MA, Branion-Calles M. Changes in commute mode
attributed to COVID-19 risk in canadian national survey data. Findings.
Network Design Lab; 2021;
doi:\href{https://doi.org/10.32866/001c.19088}{10.32866/001c.19088}}

\leavevmode\hypertarget{ref-Molloy2020Tracing}{}%
\CSLLeftMargin{35. }
\CSLRightInline{Molloy J, Tchervenkov C, Hintermann B, Axhausen KW.
Tracing the sars-CoV-2 impact: The first month in switzerland. Findings.
Network Design Lab; 2020;
doi:\href{https://doi.org/10.32866/001c.12903}{10.32866/001c.12903}}

\leavevmode\hypertarget{ref-Feyman2020effectiveness}{}%
\CSLLeftMargin{36. }
\CSLRightInline{Feyman Y, Bor J, Raifman J, Griffith KN. Effectiveness
of COVID-19 shelter-in-place orders varied by state. PLOS ONE. 2020;15:
e0245008.
doi:\href{https://doi.org/10.1371/journal.pone.0245008}{10.1371/journal.pone.0245008}}

\leavevmode\hypertarget{ref-Praharaj2020Using}{}%
\CSLLeftMargin{37. }
\CSLRightInline{Praharaj S, King D, Pettit C, Wentz E. Using aggregated
mobility data to measure the effect of COVID-19 policies on mobility
changes in sydney, london, phoenix, and pune. Findings. Network Design
Lab; 2020;
doi:\href{https://doi.org/10.32866/001c.17590}{10.32866/001c.17590}}

\leavevmode\hypertarget{ref-Paez2020using}{}%
\CSLLeftMargin{38. }
\CSLRightInline{Paez A. Using google community mobility reports to
investigate the incidence of COVID-19 in the united states. Findings.
2020; doi:\url{https://doi.org/10.32866/001c.12976}}

\leavevmode\hypertarget{ref-Noland2021mobility}{}%
\CSLLeftMargin{39. }
\CSLRightInline{Noland RB. Mobility and the effective reproduction rate
of COVID-19. Journal of Transport \& Health. 2021;20: 101016.
doi:\url{https://doi.org/10.1016/j.jth.2021.101016}}

\leavevmode\hypertarget{ref-Bhadra2021impact}{}%
\CSLLeftMargin{40. }
\CSLRightInline{Bhadra A, Mukherjee A, Sarkar K. Impact of population
density on covid-19 infected and mortality rate in india. Modeling Earth
Systems and Environment. 2021;7: 623--629.
doi:\href{https://doi.org/10.1007/s40808-020-00984-7}{10.1007/s40808-020-00984-7}}

\leavevmode\hypertarget{ref-Kadi2020population}{}%
\CSLLeftMargin{41. }
\CSLRightInline{Kadi N, Khelfaoui M. Population density, a factor in the
spread of COVID-19 in algeria: Statistic study. Bulletin of the National
Research Centre. 2020;44.
doi:\href{https://doi.org/10.1186/s42269-020-00393-x}{10.1186/s42269-020-00393-x}}

\leavevmode\hypertarget{ref-Pequeno2020air}{}%
\CSLLeftMargin{42. }
\CSLRightInline{Pequeno P, Mendel B, Rosa C, Bosholn M, Souza JL,
Baccaro F, et al. Air transportation, population density and temperature
predict the spread of COVID-19 in brazil. PeerJ. 2020;8: e9322.
doi:\href{https://doi.org/10.7717/peerj.9322}{10.7717/peerj.9322}}

\leavevmode\hypertarget{ref-Fielding2020social}{}%
\CSLLeftMargin{43. }
\CSLRightInline{Fielding-Miller RK, Sundaram ME, Brouwer K. Social
determinants of COVID-19 mortality at the county level. PLOS ONE.
2020;15: e0240151.
doi:\href{https://doi.org/10.1371/journal.pone.0240151}{10.1371/journal.pone.0240151}}

\leavevmode\hypertarget{ref-Roy2020factors}{}%
\CSLLeftMargin{44. }
\CSLRightInline{Roy S, Ghosh P. Factors affecting COVID-19 infected and
death rates inform lockdown-related policymaking. PLOS ONE. 2020;15:
e0241165.
doi:\href{https://doi.org/10.1371/journal.pone.0241165}{10.1371/journal.pone.0241165}}

\leavevmode\hypertarget{ref-Wong2020spreading}{}%
\CSLLeftMargin{45. }
\CSLRightInline{Wong DWS, Li Y. Spreading of COVID-19: Density matters.
PLOS ONE. 2020;15: e0242398.
doi:\href{https://doi.org/10.1371/journal.pone.0242398}{10.1371/journal.pone.0242398}}

\leavevmode\hypertarget{ref-Sun2020impacts}{}%
\CSLLeftMargin{46. }
\CSLRightInline{Sun Z, Zhang H, Yang Y, Wan H, Wang Y. Impacts of
geographic factors and population density on the COVID-19 spreading
under the lockdown policies of china. Science of The Total Environment.
2020;746: 141347.
doi:\href{https://doi.org/10.1016/j.scitotenv.2020.141347}{10.1016/j.scitotenv.2020.141347}}

\leavevmode\hypertarget{ref-Skorka2020macroecology}{}%
\CSLLeftMargin{47. }
\CSLRightInline{Skórka P, Grzywacz B, Moroń D, Lenda M. The macroecology
of the COVID-19 pandemic in the anthropocene. PLOS ONE. 2020;15:
e0236856.
doi:\href{https://doi.org/10.1371/journal.pone.0236856}{10.1371/journal.pone.0236856}}

\leavevmode\hypertarget{ref-White2020state}{}%
\CSLLeftMargin{48. }
\CSLRightInline{White ER, Hébert-Dufresne L. State-level variation of
initial COVID-19 dynamics in the united states. PLOS ONE. 2020;15:
e0240648.
doi:\href{https://doi.org/10.1371/journal.pone.0240648}{10.1371/journal.pone.0240648}}

\leavevmode\hypertarget{ref-Khavarian2021high}{}%
\CSLLeftMargin{49. }
\CSLRightInline{Khavarian-Garmsir AR, Sharifi A, Moradpour N. Are
high-density districts more vulnerable to the COVID-19 pandemic?
Sustainable Cities and Society. 2021;70: 102911.
doi:\href{https://doi.org/10.1016/j.scs.2021.102911}{10.1016/j.scs.2021.102911}}

\leavevmode\hypertarget{ref-Hamidi2020longitudinal}{}%
\CSLLeftMargin{50. }
\CSLRightInline{Hamidi S, Ewing R, Sabouri S. Longitudinal analyses of
the relationship between development density and the COVID-19 morbidity
and mortality rates: Early evidence from 1,165 metropolitan counties in
the united states. Health \& Place. 2020;64: 102378.
doi:\href{https://doi.org/10.1016/j.healthplace.2020.102378}{10.1016/j.healthplace.2020.102378}}

\leavevmode\hypertarget{ref-Lee2020human}{}%
\CSLLeftMargin{51. }
\CSLRightInline{Lee M, Zhao J, Sun Q, Pan Y, Zhou W, Xiong C, et al.
Human mobility trends during the early stage of the COVID-19 pandemic in
the united states. PLOS ONE. 2020;15: e0241468.
doi:\href{https://doi.org/10.1371/journal.pone.0241468}{10.1371/journal.pone.0241468}}

\leavevmode\hypertarget{ref-Amadu2021assessing}{}%
\CSLLeftMargin{52. }
\CSLRightInline{Amadu I, Ahinkorah BO, Afitiri A-R, Seidu A-A, Ameyaw
EK, Hagan JE, et al. Assessing sub-regional-specific strengths of
healthcare systems associated with COVID-19 prevalence, deaths and
recoveries in africa. PLOS ONE. 2021;16: e0247274.
doi:\href{https://doi.org/10.1371/journal.pone.0247274}{10.1371/journal.pone.0247274}}

\leavevmode\hypertarget{ref-Feng2020spread}{}%
\CSLLeftMargin{53. }
\CSLRightInline{Feng Y, Li Q, Tong X, Wang R, Zhai S, Gao C, et al.
Spatiotemporal spread pattern of the COVID-19 cases in china. PLOS ONE.
2020;15: e0244351.
doi:\href{https://doi.org/10.1371/journal.pone.0244351}{10.1371/journal.pone.0244351}}

\leavevmode\hypertarget{ref-Inbaraj2021seroprevalence}{}%
\CSLLeftMargin{54. }
\CSLRightInline{Inbaraj LR, George CE, Chandrasingh S. Seroprevalence of
COVID-19 infection in a rural district of south india: A
population-based seroepidemiological study. PLOS ONE. 2021;16: e0249247.
doi:\href{https://doi.org/10.1371/journal.pone.0249247}{10.1371/journal.pone.0249247}}

\leavevmode\hypertarget{ref-Souris2020covid}{}%
\CSLLeftMargin{55. }
\CSLRightInline{Souris M, Gonzalez J-P. COVID-19: Spatial analysis of
hospital case-fatality rate in france. PLOS ONE. 2020;15: e0243606.
doi:\href{https://doi.org/10.1371/journal.pone.0243606}{10.1371/journal.pone.0243606}}

\leavevmode\hypertarget{ref-Ahmad2020association}{}%
\CSLLeftMargin{56. }
\CSLRightInline{Ahmad K, Erqou S, Shah N, Nazir U, Morrison AR,
Choudhary G, et al. Association of poor housing conditions with COVID-19
incidence and mortality across US counties. PLOS ONE. 2020;15: e0241327.
doi:\href{https://doi.org/10.1371/journal.pone.0241327}{10.1371/journal.pone.0241327}}

\leavevmode\hypertarget{ref-Noury2021how}{}%
\CSLLeftMargin{57. }
\CSLRightInline{Noury A, François A, Gergaud O, Garel A. How does
COVID-19 affect electoral participation? Evidence from the french
municipal elections. PLOS ONE. 2021;16: e0247026.
doi:\href{https://doi.org/10.1371/journal.pone.0247026}{10.1371/journal.pone.0247026}}

\leavevmode\hypertarget{ref-Wang2021transmission}{}%
\CSLLeftMargin{58. }
\CSLRightInline{Wang F, Tan Z, Yu Z, Yao S, Guo C. Transmission and
control pressure analysis of the COVID-19 epidemic situation using
multisource spatio-temporal big data. PLOS ONE. 2021;16: e0249145.
doi:\href{https://doi.org/10.1371/journal.pone.0249145}{10.1371/journal.pone.0249145}}

\leavevmode\hypertarget{ref-Brandtner2021creatures}{}%
\CSLLeftMargin{59. }
\CSLRightInline{Brandtner C, Bettencourt LMA, Berman MG, Stier AJ.
Creatures of the state? Metropolitan counties compensated for state
inaction in initial u.s. Response to COVID-19 pandemic. PLOS ONE.
2021;16: e0246249.
doi:\href{https://doi.org/10.1371/journal.pone.0246249}{10.1371/journal.pone.0246249}}

\leavevmode\hypertarget{ref-Stephens2021impact}{}%
\CSLLeftMargin{60. }
\CSLRightInline{Stephens KE, Chernyavskiy P, Bruns DR. Impact of
altitude on COVID-19 infection and death in the united states: A
modeling and observational study. PLOS ONE. 2021;16: e0245055.
doi:\href{https://doi.org/10.1371/journal.pone.0245055}{10.1371/journal.pone.0245055}}

>>>>

A plausible behavioral adaptation during a pandemic, especially one
broadcast as widely and intensely as COVID-19, is risk compensation.
Risk compensation is a process whereby people adjust their behavior in
response to their \emph{perception} of risk {[}28--30{]}. In the case of
COVID-19, Chauhan et al. {[}31{]} have found that perception of risks in
the US varies between rural, suburban, and urban residents, with rural
residents in general displaying less concerns about the virus. It is
possible that people who listened to the message of leaders saying that
they were safe because of low density may not have taken adequate
precautions against the virus. People in dense places who could more
directly observe the impact of the pandemic may have become overly
cautious. Both Paez et al. {[}32{]} and Hamidi et al. {[}33{]} posit
this mechanism (i.e., greater compliance with social distancing in
denser regions) to explain the results of their analyses. The evidence
available does indeed show that there were important changes in behavior
during the pandemic, at least with respect to mobility {[}34--36{]};
furthermore, shelter in place orders may have had greater buy-in from
the public in higher density regions {[}37{]}, and the behavior may have
persisted beyond the duration of official social-distancing policies
{[}38{]}. In addition, there is evidence that changes in mobility
correlated with the trajectory of the pandemic {[}39,40{]}. Given the
potential for behavioral adaptation, the question of density becomes
more nuanced: it is not just a matter of proximity, but also of human
behavior, which is better studied using population-level data and
models.

@@

On the one hand, there are studies that report positive associations
between population density and various COVID-19-related outcomes. Bhadra
{[}41{]}, for example, reported a moderate positive correlation between
the spread of COVID-19 and population density at the district level in
India, however their analysis was bivariate and did not control for
other variables, such as income. Similarly, Kadi and Khelfaoui {[}42{]}
found a positive and significant correlation between number of cases and
population density in cities in Algeria in a series of simple regression
models (i.e., without other controls). A question in these relatively
simple analyses is whether density is not a proxy for other factors.
Other studies have included controls, such as Pequeno et al. {[}43{]}, a
team that reported a positive association between density and cumulative
counts of confirmed COVID-19 cases in state capitals in Brazil after
controlling for covariates, including income, transport connectivity,
and economic status. In a similar vein, Fielding-Miller et al. {[}44{]}
reported a positive relationship between the absolute number of COVID-19
deaths and population density (rate) in rural counties in the US. Roy
and Ghosh {[}45{]} used a battery of machine learning techniques to find
discriminatory factors, and a positive and significant association
between COVID-19 infection and death rates in US states. Wong and Li
{[}46{]} also found a positive and significant association between
population density and number of confirmed COVID-19 cases in US
counties, using both univariate and multivariate regressions with
spatial effects. Most recently, Sy et al. {[}19{]} reported that the
basic reproductive number of COVID-19 in US counties tends to increase
with population density, but at a decreasing rate at higher densities.

On the gripping hand, a number of studies report non-significant or
negative associations between population density and COVID-19 outcomes.
This includes the research of Sun et al. {[}47{]} who did not find
evidence of significant correlation between population density and
confirmed number of cases per day \emph{in conditions of lockdown} in
China. This finding echoes the results of Paez et al. {[}32{]}, who in
their study of provinces in Spain reported non-significant associations
between population density and infection rates in the early days of the
first wave of COVID-19, and negative significant associations in the
later part of the first lockdown. Similarly, {[}48{]} found zero or
negative associations between population density and infection
numbers/deaths by country. Fielding-Miller et al. {[}44{]} found a
negative relationship between COVID-19 deaths and population density
urban counties in the US. In their investigation of doubling time, White
and Hébert-Dufresne {[}49{]} identified a negative and significant
correlation between population density and doubling time in US states.
Likewise, {[}50{]} fond a small negative (and significant) association
between population density and COVID-19 morbidity in districts in
Tehran. And two of the most complete studies in the US by Hamidi et al.
{[}33,51{]} used an extensive set of controls to find negative and
significant correlations between density at the level of counties in the
US and COVID-19 cases and fatalities.

@@

There are various reasons why a project can fail to be reproducible. In
some cases, there might be legitimate reasons to withhold the data,
perhaps due to confidentiality and privacy reasons {[}e.g., 52{]}. But
in many other cases the data are publicly available, as has been
commonly the case with population-level COVID-19 information. Often the
provenance of the data is documented, but the data themselves are not
shared {[}25,33,41,44,51,e.g., 53,54--56{]}. As any researcher can
attest, whether a graduate student or a seasoned scientist, collecting,
organizing, and preparing data for a project can take a substantial
amount of time. Pointing to the sources of data, even when these sources
are public, is a small step towards reproducibility-but only a very
small one. Faced with the prospect of having to recreate a data set from
raw sources is probably sufficient to dissuade all but the most
dedicated (or stubborn) researcher. This is true even if part of the
data are shared {[}e.g., 46{]}. In other cases, data are shared, but the
processes to document the preparation of the data are not fully
documented {[}48,e.g., 57{]}. These processes matter, as shown by the
errors in the spreadsheets of Reinhart and Rogoff {[}16{]}, and the data
of biologist Jonathan Pruitt that led to an ``avalanche'' of paper
retractions\footnote{\url{https://doi.org/10.1038/d41586-020-00287-y}}.
Another situation is when papers share well-documented data, but fail to
provide the code used in the analysis {[}43,e.g., 58,59{]}. Making code
available only ``on demand'' {[}e.g., 60{]} is an unnecessary barrier
when most journals offer the facility to share supplemental materials
online. Then there are those papers that strive towards reproducibility,
sharing well-documented processes and data, as well as the code used in
any analyses reported {[}e.g., 32,37,49,61{]}.

@@

\leavevmode\hypertarget{ref-Chauhan2021covid}{}%
\CSLLeftMargin{31. }
\CSLRightInline{Chauhan RS, Capasso da Silva D, Salon D, Shamshiripour
A, Rahimi E, Sutradhar U, et al. COVID-19 related attitudes and risk
perceptions across urban, rural, and suburban areas in the united
states. Findings. Network Design Lab; 2021;
doi:\href{https://doi.org/10.32866/001c.23714}{10.32866/001c.23714}}

\leavevmode\hypertarget{ref-Paez2020spatio}{}%
\CSLLeftMargin{32. }
\CSLRightInline{Paez A, Lopez FA, Menezes T, Cavalcanti R, Pitta MG da
R. A spatio-temporal analysis of the environmental correlates of
COVID-19 incidence in spain. Geographical Analysis. 2020;n/a.
doi:\href{https://doi.org/10.1111/gean.12241}{10.1111/gean.12241}}

\leavevmode\hypertarget{ref-Hamidi2020density}{}%
\CSLLeftMargin{33. }
\CSLRightInline{Hamidi S, Sabouri S, Ewing R. Does density aggravate the
COVID-19 pandemic? Journal of the American Planning Association.
2020;86: 495--509.
doi:\href{https://doi.org/10.1080/01944363.2020.1777891}{10.1080/01944363.2020.1777891}}

\leavevmode\hypertarget{ref-Jamal2020Changes}{}%
\CSLLeftMargin{34. }
\CSLRightInline{Jamal S, Paez A. Changes in trip-making frequency by
mode during COVID-19. Findings. Network Design Lab; 2020;
doi:\href{https://doi.org/10.32866/001c.17977}{10.32866/001c.17977}}

\leavevmode\hypertarget{ref-Harris2021Changes}{}%
\CSLLeftMargin{35. }
\CSLRightInline{Harris MA, Branion-Calles M. Changes in commute mode
attributed to COVID-19 risk in canadian national survey data. Findings.
Network Design Lab; 2021;
doi:\href{https://doi.org/10.32866/001c.19088}{10.32866/001c.19088}}

\leavevmode\hypertarget{ref-Molloy2020Tracing}{}%
\CSLLeftMargin{36. }
\CSLRightInline{Molloy J, Tchervenkov C, Hintermann B, Axhausen KW.
Tracing the sars-CoV-2 impact: The first month in switzerland. Findings.
Network Design Lab; 2020;
doi:\href{https://doi.org/10.32866/001c.12903}{10.32866/001c.12903}}

\leavevmode\hypertarget{ref-Feyman2020effectiveness}{}%
\CSLLeftMargin{37. }
\CSLRightInline{Feyman Y, Bor J, Raifman J, Griffith KN. Effectiveness
of COVID-19 shelter-in-place orders varied by state. PLOS ONE. 2020;15:
e0245008.
doi:\href{https://doi.org/10.1371/journal.pone.0245008}{10.1371/journal.pone.0245008}}

\leavevmode\hypertarget{ref-Praharaj2020Using}{}%
\CSLLeftMargin{38. }
\CSLRightInline{Praharaj S, King D, Pettit C, Wentz E. Using aggregated
mobility data to measure the effect of COVID-19 policies on mobility
changes in sydney, london, phoenix, and pune. Findings. Network Design
Lab; 2020;
doi:\href{https://doi.org/10.32866/001c.17590}{10.32866/001c.17590}}

\leavevmode\hypertarget{ref-Paez2020using}{}%
\CSLLeftMargin{39. }
\CSLRightInline{Paez A. Using google community mobility reports to
investigate the incidence of COVID-19 in the united states. Findings.
2020; doi:\url{https://doi.org/10.32866/001c.12976}}

\leavevmode\hypertarget{ref-Noland2021mobility}{}%
\CSLLeftMargin{40. }
\CSLRightInline{Noland RB. Mobility and the effective reproduction rate
of COVID-19. Journal of Transport \& Health. 2021;20: 101016.
doi:\url{https://doi.org/10.1016/j.jth.2021.101016}}

\leavevmode\hypertarget{ref-Bhadra2021impact}{}%
\CSLLeftMargin{41. }
\CSLRightInline{Bhadra A, Mukherjee A, Sarkar K. Impact of population
density on covid-19 infected and mortality rate in india. Modeling Earth
Systems and Environment. 2021;7: 623--629.
doi:\href{https://doi.org/10.1007/s40808-020-00984-7}{10.1007/s40808-020-00984-7}}

\leavevmode\hypertarget{ref-Kadi2020population}{}%
\CSLLeftMargin{42. }
\CSLRightInline{Kadi N, Khelfaoui M. Population density, a factor in the
spread of COVID-19 in algeria: Statistic study. Bulletin of the National
Research Centre. 2020;44.
doi:\href{https://doi.org/10.1186/s42269-020-00393-x}{10.1186/s42269-020-00393-x}}

\leavevmode\hypertarget{ref-Pequeno2020air}{}%
\CSLLeftMargin{43. }
\CSLRightInline{Pequeno P, Mendel B, Rosa C, Bosholn M, Souza JL,
Baccaro F, et al. Air transportation, population density and temperature
predict the spread of COVID-19 in brazil. PeerJ. 2020;8: e9322.
doi:\href{https://doi.org/10.7717/peerj.9322}{10.7717/peerj.9322}}

\leavevmode\hypertarget{ref-Fielding2020social}{}%
\CSLLeftMargin{44. }
\CSLRightInline{Fielding-Miller RK, Sundaram ME, Brouwer K. Social
determinants of COVID-19 mortality at the county level. PLOS ONE.
2020;15: e0240151.
doi:\href{https://doi.org/10.1371/journal.pone.0240151}{10.1371/journal.pone.0240151}}

\leavevmode\hypertarget{ref-Roy2020factors}{}%
\CSLLeftMargin{45. }
\CSLRightInline{Roy S, Ghosh P. Factors affecting COVID-19 infected and
death rates inform lockdown-related policymaking. PLOS ONE. 2020;15:
e0241165.
doi:\href{https://doi.org/10.1371/journal.pone.0241165}{10.1371/journal.pone.0241165}}

\leavevmode\hypertarget{ref-Wong2020spreading}{}%
\CSLLeftMargin{46. }
\CSLRightInline{Wong DWS, Li Y. Spreading of COVID-19: Density matters.
PLOS ONE. 2020;15: e0242398.
doi:\href{https://doi.org/10.1371/journal.pone.0242398}{10.1371/journal.pone.0242398}}

\leavevmode\hypertarget{ref-Sun2020impacts}{}%
\CSLLeftMargin{47. }
\CSLRightInline{Sun Z, Zhang H, Yang Y, Wan H, Wang Y. Impacts of
geographic factors and population density on the COVID-19 spreading
under the lockdown policies of china. Science of The Total Environment.
2020;746: 141347.
doi:\href{https://doi.org/10.1016/j.scitotenv.2020.141347}{10.1016/j.scitotenv.2020.141347}}

\leavevmode\hypertarget{ref-Skorka2020macroecology}{}%
\CSLLeftMargin{48. }
\CSLRightInline{Skórka P, Grzywacz B, Moroń D, Lenda M. The macroecology
of the COVID-19 pandemic in the anthropocene. PLOS ONE. 2020;15:
e0236856.
doi:\href{https://doi.org/10.1371/journal.pone.0236856}{10.1371/journal.pone.0236856}}

\leavevmode\hypertarget{ref-White2020state}{}%
\CSLLeftMargin{49. }
\CSLRightInline{White ER, Hébert-Dufresne L. State-level variation of
initial COVID-19 dynamics in the united states. PLOS ONE. 2020;15:
e0240648.
doi:\href{https://doi.org/10.1371/journal.pone.0240648}{10.1371/journal.pone.0240648}}

\leavevmode\hypertarget{ref-Khavarian2021high}{}%
\CSLLeftMargin{50. }
\CSLRightInline{Khavarian-Garmsir AR, Sharifi A, Moradpour N. Are
high-density districts more vulnerable to the COVID-19 pandemic?
Sustainable Cities and Society. 2021;70: 102911.
doi:\href{https://doi.org/10.1016/j.scs.2021.102911}{10.1016/j.scs.2021.102911}}

\leavevmode\hypertarget{ref-Hamidi2020longitudinal}{}%
\CSLLeftMargin{51. }
\CSLRightInline{Hamidi S, Ewing R, Sabouri S. Longitudinal analyses of
the relationship between development density and the COVID-19 morbidity
and mortality rates: Early evidence from 1,165 metropolitan counties in
the united states. Health \& Place. 2020;64: 102378.
doi:\href{https://doi.org/10.1016/j.healthplace.2020.102378}{10.1016/j.healthplace.2020.102378}}

\leavevmode\hypertarget{ref-Lee2020human}{}%
\CSLLeftMargin{52. }
\CSLRightInline{Lee M, Zhao J, Sun Q, Pan Y, Zhou W, Xiong C, et al.
Human mobility trends during the early stage of the COVID-19 pandemic in
the united states. PLOS ONE. 2020;15: e0241468.
doi:\href{https://doi.org/10.1371/journal.pone.0241468}{10.1371/journal.pone.0241468}}

\leavevmode\hypertarget{ref-Amadu2021assessing}{}%
\CSLLeftMargin{53. }
\CSLRightInline{Amadu I, Ahinkorah BO, Afitiri A-R, Seidu A-A, Ameyaw
EK, Hagan JE, et al. Assessing sub-regional-specific strengths of
healthcare systems associated with COVID-19 prevalence, deaths and
recoveries in africa. PLOS ONE. 2021;16: e0247274.
doi:\href{https://doi.org/10.1371/journal.pone.0247274}{10.1371/journal.pone.0247274}}

\leavevmode\hypertarget{ref-Feng2020spread}{}%
\CSLLeftMargin{54. }
\CSLRightInline{Feng Y, Li Q, Tong X, Wang R, Zhai S, Gao C, et al.
Spatiotemporal spread pattern of the COVID-19 cases in china. PLOS ONE.
2020;15: e0244351.
doi:\href{https://doi.org/10.1371/journal.pone.0244351}{10.1371/journal.pone.0244351}}

\leavevmode\hypertarget{ref-Inbaraj2021seroprevalence}{}%
\CSLLeftMargin{55. }
\CSLRightInline{Inbaraj LR, George CE, Chandrasingh S. Seroprevalence of
COVID-19 infection in a rural district of south india: A
population-based seroepidemiological study. PLOS ONE. 2021;16: e0249247.
doi:\href{https://doi.org/10.1371/journal.pone.0249247}{10.1371/journal.pone.0249247}}

\leavevmode\hypertarget{ref-Souris2020covid}{}%
\CSLLeftMargin{56. }
\CSLRightInline{Souris M, Gonzalez J-P. COVID-19: Spatial analysis of
hospital case-fatality rate in france. PLOS ONE. 2020;15: e0243606.
doi:\href{https://doi.org/10.1371/journal.pone.0243606}{10.1371/journal.pone.0243606}}

\leavevmode\hypertarget{ref-Ahmad2020association}{}%
\CSLLeftMargin{57. }
\CSLRightInline{Ahmad K, Erqou S, Shah N, Nazir U, Morrison AR,
Choudhary G, et al. Association of poor housing conditions with COVID-19
incidence and mortality across US counties. PLOS ONE. 2020;15: e0241327.
doi:\href{https://doi.org/10.1371/journal.pone.0241327}{10.1371/journal.pone.0241327}}

\leavevmode\hypertarget{ref-Noury2021how}{}%
\CSLLeftMargin{58. }
\CSLRightInline{Noury A, François A, Gergaud O, Garel A. How does
COVID-19 affect electoral participation? Evidence from the french
municipal elections. PLOS ONE. 2021;16: e0247026.
doi:\href{https://doi.org/10.1371/journal.pone.0247026}{10.1371/journal.pone.0247026}}

\leavevmode\hypertarget{ref-Wang2021transmission}{}%
\CSLLeftMargin{59. }
\CSLRightInline{Wang F, Tan Z, Yu Z, Yao S, Guo C. Transmission and
control pressure analysis of the COVID-19 epidemic situation using
multisource spatio-temporal big data. PLOS ONE. 2021;16: e0249145.
doi:\href{https://doi.org/10.1371/journal.pone.0249145}{10.1371/journal.pone.0249145}}

\leavevmode\hypertarget{ref-Brandtner2021creatures}{}%
\CSLLeftMargin{60. }
\CSLRightInline{Brandtner C, Bettencourt LMA, Berman MG, Stier AJ.
Creatures of the state? Metropolitan counties compensated for state
inaction in initial u.s. Response to COVID-19 pandemic. PLOS ONE.
2021;16: e0246249.
doi:\href{https://doi.org/10.1371/journal.pone.0246249}{10.1371/journal.pone.0246249}}

\leavevmode\hypertarget{ref-Stephens2021impact}{}%
\CSLLeftMargin{61. }
\CSLRightInline{Stephens KE, Chernyavskiy P, Bruns DR. Impact of
altitude on COVID-19 infection and death in the united states: A
modeling and observational study. PLOS ONE. 2021;16: e0245055.
doi:\href{https://doi.org/10.1371/journal.pone.0245055}{10.1371/journal.pone.0245055}}
